---
title: 'Part 2: chi-sqaure with lizards'' tails'
author: "Devon Rossi"
date: "11/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Read in data
```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```


Get counts of lizards by species:
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```

## Make a subset

- UTST lizards
- Sites: IBPE (enclosure grassland), RABB (rabbit mesquite site), "EAST" (east tarbush site)
- Tail condition is recorded (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))

#use unique function in console to make sure your filter is restricting the variables you want (i.e. sites)
```

## Find counts and proportions of lizards with broken or whole tails at each site

```{r}
utst %>% 
  count(site, tail)
```


```{r}
tail_count <- utst %>% 
  janitor::tabyl(site, tail) #contingency table

tail_proportions <- tail_count %>% 
  janitor::adorn_percentages() %>% #percent
  janitor::adorn_pct_formatting(digits = 2) %>% #round to 2 places
  janitor::adorn_ns() #counts
```


## Convert to a contingency table: Set up for Chi-Squared test

Use `column_to_rowname()`

```{r}
lizard_ct <- tail_count %>% 
  column_to_rownames(var = "site")

#site names (EAST, IBPE, RABB) are now listed as row names!!!
```


## Chi-Square

- Are tail breakage and site independent? (Site has no impact on tail breakage - expect similar proportions across 3 diff sites)

- Are those proportions at each site of lizards that have B and W so different that we have enough evidence to REJECT NULL HYPOTHESIS of independence which would allow us to conclude that 
yes, there is a significant association of site on tail breakage?

## Use Broom::tidy(lizard_x2) to set up for in-line referencing
```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)
```

Chi-squared value of 13.06

Degrees of freedom: number of rows(3) - 1 = 2 times
number of columns(2) - 1 = 1 

2 x 1 = 2 degrees of freedom

p-value = 0.0014

*Using a sig level of .05, yes we have significant evidence to reject the null that site and tail breakage are NOT independent!!!*

*There is a sig association b/w site and tail breakage*

There is a significant association (i.e. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`).







